---
title: "Data Exploration"
author: "Jay"
date: "June 10, 2016"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, cache.path = 'cache/', fig.path='figure/')
```


```{r libraries}
# install.packages(c('devtools', 'curl', 'corrplot', 'Cairo', 'tibble'))
# devtools::install_github("hadley/assertthat")
# devtools::install_github("rstats-db/bigrquery")

library(bigrquery)
library(curl)
library(corrplot)
library(Cairo)
library(ggplot2)
```

```{r load_data}
setwd("C:\\Dropbox\\Projects\\kaggle\\datalab-projects\\Xiaojukeji Algorithm Competition\\R")

filename <- "xjk_algo_comp.gaps.Rda"

if (file.exists(filename)) {
  load(filename)
} else {
  project <- "datalab-projects-1331" # put your projectID here
 
  sql <- 'SELECT * FROM[datalab-projects-1331:xjk_algo_comp.gaps];'
   
  data <- query_exec(sql, project=project, max_pages=Inf)
  save(data,file=filename)
}
data[is.na(data)] <- 0
```

# Corrplot
```{r corrplot}
data1 <- data[ , !names(data) %in% c("timeslot", "district_id")]
cors <- cor(data1, use="complete")
corrplot(cors, method="color", order="hclust", tl.cex = 0.8)
```
```{r corlist}
corrplot(cors[1:163,164, drop=FALSE], method="color", tl.cex = 0.5, cl.pos='n')
```

Clearer version is available at file `corrplot.png` (4000px width).

From there we see that `gap` is only strongly correlated with `sum_price`.

# Which days in week have different gaps?
```{r days_in_week}
ggplot(aes(y=gap, x=as.factor(day_in_week)), data=data) +
  geom_boxplot() +
  scale_y_log10()
```

Looks like there are usually less gaps on Sunday. Variable `is_sunday` may need to be created.

```{r corr2}
data$is_sunday <- ifelse(data$day_in_week == 0, TRUE, FALSE)
cor1 <- cor(x=data$is_sunday, y=data$gap)
cor2 <- cor(x=data$sum_price, y=data$gap)
cor3 <- cor(x=data$poi1, y=data$gap)
print(paste("correlation between is_sunday and gap: ", cor1, sep=" "))
print(paste("correlation between sum_price and gap: ", cor2, sep=" "))
print(paste("correlation between poi1 and gap: ", cor3, sep=" "))
```

As it turned out, the correlation between `is_sunday` and `gap` is rather low (around 0.02) so we can't really use it.

# Get ordered correlation
Useful for feature selection step.

```{r}
cors <- cor(x=data$gap, y=data1)
corsdf <- as.data.frame(as.table(cors))
corsdf=corsdf[order(-abs(corsdf$Freq)),]
x <- subset(corsdf, abs(Freq) > 0.3)$Var2
print(paste(x, collapse=", "))
```

# Traffic
How do gaps differ on different traffic levels?

```{r traffic1}
ggplot(aes(y=gap, x=tj_level1_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

After 3500 `traffic_tj_level1` some high gaps are starting to show, but their numbers are insignificant compared to the entire general trend.

Let's see if log scaling `gap` variable will unravel more information.

```{r traffic2}
ggplot(aes(y=gap, x=tj_level1_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  scale_y_log10() +
  geom_smooth()
```

It doesn't look so, let's move on.

# Previous slot's gap

Maybe previous slot's gap affects next gap, since they are only separated by 10 minutes. Let's explore this.

```{r prevgap}
ggplot(aes(y=gap, x=gap_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

Bingo! `gap` and `previous_gap` are highly correlated it seems.


# Weather

Weather data have really low correlation with `gap`, but I wonder if there are some hidden structures we can find here.

```{r weather1}
ggplot(aes(y=gap, x=as.factor(weather_1_slots_ago)), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

It looks like weather 1 has several high gaps. Let's see the distributions more clearly by using box plot and log-scale the y-axis.

```{r weather1a}
ggplot(aes(y=gap, x=as.factor(weather_1_slots_ago)), data=data) +
  geom_boxplot() +
  geom_smooth() +
  scale_y_log10()
```

There doesn't seem to be any prevalent trend here.

```{r weather2}
ggplot(aes(y=gap, x=pm25_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

```{r weather2a}
ggplot(aes(y=gap, x=pm25_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth() +
  scale_y_log10()
```

```{r weather3}
ggplot(aes(y=gap, x=as.factor(temperature_1_slots_ago)), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

```{r weather3a}
ggplot(aes(y=gap, x=as.factor(temperature_1_slots_ago)), data=data) +
  geom_boxplot() +
  geom_smooth() +
  scale_y_log10()
```

It seems like there are three clusters of weather temperatures that correspond to different level of ``

```{r weather4}
ggplot(aes(y=weather_pm25, x=as.factor(temperature_1_slots_ago), color=gap), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

# Timeofday_slot with gap
```{r timeofday_slot}
ggplot(aes(y=gap, x=as.factor(timeofday_slot)), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

Looks like there are some hours where gaps are surging, let's try to plot it clearer

```{r timeofday_slot}
ggplot(aes(y=gap, x=as.factor(timeofday_slot)), data=data) +
  geom_boxplot() +
  scale_y_log10()
```

Surge happens at min 50 to 53.

# Gap correlations

```{r gap_cor_1}
ggplot(aes(y=gap, x=gap_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  scale_y_log10()
```

```{r gap_cor_1}
ggplot(aes(y=gap, x=gap_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```
```{r gap_cor_2}
ggplot(aes(y=gap, x=gap_2_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```
```{r gap_cor_3}
ggplot(aes(y=gap, x=gap_3_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

```{r gap_sum_cor_1}
ggplot(aes(y=gap, x=sum_price_1_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

```{r gap_sum_cor_2}
ggplot(aes(y=gap, x=sum_price_2_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```

```{r gap_sum_cor_3}
ggplot(aes(y=gap, x=sum_price_3_slots_ago), data=data) +
  geom_jitter(alpha=0.4) +
  geom_smooth()
```